---
title: "Advanced REDCap API & RStudio Exercise"
author: "LV"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: united
    highlight: tango
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

## Prerequisites 🔧

```{r prerequisites}
# Install required packages
install.packages("redcapAPI")    # CRAN package for REDCap API
install.packages("tidyverse")    # Collection of data science packages
```

- **R & RStudio installed**  
- **REDCap project URL** and **API token** obtained from your REDCap administrator  
- Internet access to your REDCap server  

## 1. Connect to REDCap API 🔑

```{r connect}
library(redcapAPI)
library(tidyverse)

redcap_url <- "https://redcap.example.edu/api/"   # Your REDCap API endpoint
api_token  <- "YOUR_API_TOKEN_HERE"               # Keep this secret!

rccon <- redcapConnection(
  url   = redcap_url,
  token = api_token
)
```

## 2. Export Project Metadata 🗂️

```{r export_metadata}
dict <- exportMetaData(rccon)      

dict %>% 
  select(field_name, form_name, field_type) %>% 
  head()
```

## 3. Export Records 📥

```{r export_records}
records <- exportRecords(
  rccon,
  factors = FALSE,
  labels  = TRUE
)

glimpse(records)
```

## 4. Data Cleaning & Summary with tidyverse 📊

```{r data_cleaning}
records2 <- records %>% 
  mutate(
    height_m = height_cm / 100,
    bmi      = weight_kg / (height_m^2)
  ) %>% 
  select(record_id, age, height_cm, weight_kg, bmi)

records2 %>% 
  summarise(
    n     = n(),
    mean  = mean(bmi, na.rm = TRUE),
    sd    = sd(bmi, na.rm = TRUE),
    max   = max(bmi, na.rm = TRUE),
    min   = min(bmi, na.rm = TRUE)
  )
```

## 5. Import Modified Records ↩️

```{r import_records}
to_update <- records2 %>% 
  filter(bmi > 30) %>% 
  transmute(
    record_id,
    high_bmi_flag = 1
  )

importResults <- importRecords(
  rccon,
  to_update
)

importResults
```

## 6. Error Handling & Logging 🐞

```{r error_handling}
records_safe <- tryCatch(
  exportRecords(rccon),
  error = function(e) {
    message("Error exporting records: ", e$message)
    tibble()
  }
)

# (Optional) If supported by your REDCap version:
# api_log <- showUsage(rccon)
# print(api_log)
```

## Next Steps & Best Practices 🚀

- **Secure token storage:** use environment variables (`Sys.getenv()`) or a keyring.  
- **Automate workflows:** schedule R scripts or RMarkdown renderings.  
- **Version control metadata:** commit the `dict` CSV to Git to track instrument changes.  
- **Explore other endpoints:** files, participants, user rights, survey invitations, etc.  

```{r}
# End of exercise
```
